### 2020-08-20 response/response-demography-3.tex

\item Perhaps the conclusion could start with a brief overview of your
topic and research question again to remind the reader of the overall
motivation for the study.

[Response:]

I have added a brief synopsis of the changes in India at the beginning of the Conclusion. 
I recap the three questions after that. 
I have also edited the Conclusion to tie better with the questions.

\item I see the point to reviewer’s comment \#4 about the discussion on
women’s labor supply. It’s a complex topic to introduce in a few brief
sentences. Similarly, referring to Goldin’s U-shaped theory without a
more detailed explanation could be confusing for those not well versed
in this topic. Perhaps you could also mention that decreasing FLFP is
unexpected for developing countries and some potential reasons for this
in India.

@@

I believe the female labor supply discussion was disjointed and unclear because I had split the 
background and conceptual framework sections, making it diﬀicult to see the connection. 
As explained above, I have combined and reorganized those two sections into one and removed 
the Goldin theory discussion since it is not directly relevant to my main argument.

\item Overall, I felt that the paper is still quite densely written and
it is not easy to follow what is going on. Considerable time needs to be
spent re-writing the paper and making it easier to read.

[Response:]

As I have outlined above, I have reorganized and significantly rewritten large parts of 
the paper to make it easier to follow. 
I have also had an editor go through the paper to help with grammar and clarity.

### 2020-08-20 paper/sex_selection_spacing-ver4.tex
Editing (#26)
* Editing and update with Demography style guide

- Edit based on proofreader comments
- Consistent presentation across appendix and main paper

* Edit Abstract and Introduction

- Focus on main questions

* Edit Background and Conceptual Framework section

*  Edit Empirical Strategy section

- Remove average discussion of Estimation Strategy section
- Fix percentile examples
- Remove unneeded formula

* Edit Data section

* Major editing of birth interval section(s) 

- Combined percentile spacing, sex ratio, parity progression graphs
- Thinner lines for easier separating lines

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_20} 
    \end{minipage}
}
\caption{Percentage of married women who were working at the time of the 
survey by age group and area of residence}
\label{fig:work_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_20} 
    \end{minipage}
}
\caption{Percentage of women paid cash or cash and in-kind of those women who were 
working at the time of the survey by age group and area of residence}
\label{fig:work_cash_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_20} 
    \end{minipage}
}
\caption{Percentage women who worked for a family member of those 
working at the time of the survey, by age group and area of residence}
\label{fig:work_family_by_survey}
\end{figure}

@@

Interpretation of the model coefficients is challenging \citep{thomas96}.
It is, however, possible to calculate the predicted probabilities of 
having a boy, $b$, and of having a girl, $g$, in period $t$, conditional on 
a set of explanatory variables and not having had a child before that period, as
\begin{align}
P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt}, t ) 
& =  
\frac{ \exp(D_j(t) + \alpha_{1t}' \mathbf{Z}_{kt} + \beta_1' \mathbf{X}_{k} )}
{1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt} ' \mathbf{Z}_{kt} + \beta_l ' \mathbf{X}_{k})}
\label{eq:probability_boy} \\
P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t ) 
& =  
\frac{ \exp(D_j(t) + \alpha_{2t}'\mathbf{Z}_{kt} + \beta_2'\mathbf{X}_{k} )}
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{kt} + \beta_l'\mathbf{X}_{k})}
\label{eq:probability_girl}
\end{align}
It is then straightforward to calculate the estimated percentage of children born that 
are boys, $\hat{Y}$, at each $t$:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100.
\label{eq:probability_son}
\end{equation}
Combining the percentage boys and the likelihood of exiting the spell 
across all $t$ gives the predicted percent boys born over the entire spell.%
\footnote{
Imagine $T=2$. 
If 54 and 66 percent of births are boys and the likelihood of giving birth 20 and 40 percent, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent boys. 
}

@@

The ``first-born'' sex ratios illustrate the systematic recall error problem well.
In all four surveys around 55 percent of children reported as first-born are boys
for the first cohort of births observed.
Given that these cohorts cover from 1960-1964 to 1980-1984, which is before sex selection 
techniques became available in India, the most likely explanation for the skewed sex ratio 
is that some children listed as first-borns were not, in fact, the first children born in 
their families.
Instead, for a substantial proportion of families, their first-born was a girl who died 
and went unreported when enumerators asked about birth history.

@@

The relationship between the length of marriage and recall error can also be seen in 
Figures \ref{fig:sex_ratio_recall_rounds_bo1} and \ref{fig:sex_ratio_recall_rounds_bo2}, 
which show the observed sex ratio for children reported as first born as a function of 
the duration of marriage at the time of the survey.
The solid line is the sex ratio of children reported as first-born by the number of years 
between the survey and marriage, while the dashed lines indicate the 95 percent confidence 
interval and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes I group years into twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys is increasingly above the expected 
value the longer ago the parents were married.

@@

Figures \ref{fig:spacing_low_urban} and \ref{fig:spacing_highest_rural} show 25th, 50th,
and 75th percentile birth intervals, the sex ratio, and the probability of parity
progression by spell for urban women with no education and rural women with twelve or
more years of education, respectively.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{bs_highest_rural}
\caption{Percentile birth intervals, sex ratios, and parity progression  
for rural women with twelve or more years of education by spell, sex composition, and period}
\label{fig:spacing_highest_rural}
\end{figure}

@@

The second set of tables, 
Tables \ref{tab:avg_sex_ratio_low}, \ref{tab:avg_sex_ratio_med}, 
\ref{tab:avg_sex_ratio_high}, and \ref{tab:avg_sex_ratio_highest}, show predicted average 
birth intervals, sex ratios, and probabilities of having a birth by decade, spell, and sex 
composition for the four education levels separated by the area of residence, together 
with bootstrapped standard errors for all three outcomes.
To find the average birth interval, I calculate, for each woman, the probability of 
giving birth in each t, and her expected spell length from these probabilities. 
I then average the individual expected spell lengths across women using their parity 
progression probabilities as weights. Finally, I add nine months because spells begin 
nine months after the previous birth.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{mortality_spell_3_low_med}
\caption{Infant mortality by preceding birth interval across periods for third child of women with 
no education and women with one to seven years of education}
\label{fig:mortality_low_med_spell_3}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{mortality_spell_3_high_highest}
\caption{Infant mortality by preceding birth interval across periods for third child of women with 
eight to eleven and twelve and above years of education}
\label{fig:mortality_high_highest_spell_3}
\end{figure}

>>>>

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30--39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20--29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_20} 
    \end{minipage}
}
\caption{Percentage of married women who were working at the time of the 
survey by age group and area of residence}
\label{fig:work_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30--39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20--29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_20} 
    \end{minipage}
}
\caption{Percentage of women paid cash or cash and in-kind of those women who were 
working at the time of the survey by age group and area of residence}
\label{fig:work_cash_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30--39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20--29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_20} 
    \end{minipage}
}
\caption{Percentage women who worked for a family member of those 
working at the time of the survey, by age group and area of residence}
\label{fig:work_family_by_survey}
\end{figure}

@@

Interpretation of the model coefficients is challenging \citep{thomas96}.
It is, however, possible to calculate the predicted probabilities of 
having a boy, $b$, and of having a girl, $g$, in period $t$, conditional on 
a set of explanatory variables and not having had a child before that period, as
\begin{align}
P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt}, t ) 
& =  
\frac{ \exp(D_j(t) + \alpha_{1t}' \mathbf{Z}_{kt} + \beta_1' \mathbf{X}_{k} )}
{1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt} ' \mathbf{Z}_{kt} + \beta_l ' \mathbf{X}_{k})}
\label{eq:probability_boy} \\
P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t ) 
& =  
\frac{ \exp(D_j(t) + \alpha_{2t}'\mathbf{Z}_{kt} + \beta_2'\mathbf{X}_{k} )}
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{kt} + \beta_l'\mathbf{X}_{k})}
\label{eq:probability_girl}
\end{align}
It is then straightforward to calculate the estimated percentage of children born that 
are boys, $\hat{Y}$, at each $t$:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100.
\label{eq:probability_son}
\end{equation}
Combining the percentage boys and the likelihood of exiting the spell 
across all $t$ gives the predicted percent boys born over the entire spell.%
\footnote{
Imagine $T=2$. 
If 54\% and 66\% of births are boys and the likelihood of giving birth 20\% and 40\%, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$\% boys. 
}

@@

The ``first-born'' sex ratios illustrate the systematic recall error problem well.
In all four surveys around 55\% of children reported as first-born are boys
for the first cohort of births observed.
Given that these cohorts cover from 1960-1964 to 1980-1984, which is before sex selection 
techniques became available in India, the most likely explanation for the skewed sex ratio 
is that some children listed as first-borns were not, in fact, the first children born in 
their families.
Instead, for a substantial proportion of families, their first-born was a girl who died 
and went unreported when enumerators asked about birth history.

@@

The relationship between the length of marriage and recall error can also be seen in 
Figures \ref{fig:sex_ratio_recall_rounds_bo1} and \ref{fig:sex_ratio_recall_rounds_bo2}, 
which show the observed sex ratio for children reported as first born as a function of 
the duration of marriage at the time of the survey.
The solid line is the sex ratio of children reported as first-born by the number of years 
between the survey and marriage, while the dashed lines indicate the 95\% confidence 
interval and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes I group years into twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys is increasingly above the expected 
value the longer ago the parents were married.

@@

Figures \ref{fig:spacing_low_urban} and \ref{fig:spacing_highest_rural} show 25th, 50th,
and 75th percentile birth intervals, the sex ratio, and the probability of parity
progression by spell for urban women with no education and rural women with 12 or
more years of education, respectively.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{bs_highest_rural}
\caption{Percentile birth intervals, sex ratios, and parity progression  
for rural women with 12 or more years of education by spell, sex composition, and period}
\label{fig:spacing_highest_rural}
\end{figure}

@@

The second set of tables, 
Tables \ref{tab:avg_sex_ratio_low}, \ref{tab:avg_sex_ratio_med}, 
\ref{tab:avg_sex_ratio_high}, and \ref{tab:avg_sex_ratio_highest}, show predicted average 
birth intervals, sex ratios, and probabilities of having a birth by decade, spell, and sex 
composition for the four education levels separated by the area of residence, together 
with bootstrapped standard errors for all three outcomes.
To find the average birth interval, I calculate, for each woman, the probability of 
giving birth in each $t$, and her expected spell length from these probabilities. 
I then average the individual expected spell lengths across women using their parity 
progression probabilities as weights. Finally, I add nine months because spells begin 
nine months after the previous birth.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{mortality_spell_3_low_med}
\caption{Infant mortality by preceding birth interval across periods for third child of women with 
no education and women with 1--7 years of education}
\label{fig:mortality_low_med_spell_3}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio=true]{mortality_spell_3_high_highest}
\caption{Infant mortality by preceding birth interval across periods for third child of women with 
8--11 and 12 and above years of education}
\label{fig:mortality_high_highest_spell_3}
\end{figure}
